Group Theory in Physics (Course Note)

### 2016-10-10 Complex Geometry - index/misc/complex geometry-misc.tex
New class notes. A fews new lines in fibration (Complex Geometry)

    \subsection{Fibration}
    \label{sec:Fibration}
    \paragraph{Lift of morphisms}
    \begin{defi}[Lift of morphisms]
    \nomenclature{Lift of morphisms}{\nomrefpage.}
        The \textbf{lift} of a morphism $f: Y\to B$ along an epimorphism
        \footnote{Epimorphism is the category analogy of surjective
        functions in set theory}. (or more general map) $p:X\to B$ is a
        morphism $\tilde{f}: Y\to X$ such that $f = p\circ\tilde{f}$.
    \end{defi}
    $$
    \xymatrix{
        X\ar[r]^p                    & B \\
        Y\ar[ur]^f\ar[u]^{\tilde{f}} & \\
    }
    $$
    \begin{defi}[Lift property]
    \nomenclature{Lift property}{\nomrefpage.}
        We say that $f$ has a \textbf{left lifting property} w.r.t $g$, or
        equivalently that $g$ has a \textbf{right lifting property} w.r.t
        $f$, if and only if for every commutative diagram below:
        $$
        \xymatrix{
        a\ar[r]^u\ar[d]^f & c\ar[d]^g \\
        b\ar[r]^v         & d
        }
        $$
        there is an arrow $\gamma$, s.t. both triangles in the followinng
        diagram commutes.
        $$ \xymatrix{
        a\ar[r]^u\ar[d]^f         & c\ar[d]^g \\
        b\ar[r]^v\ar[ru]^{\gamma} & d
        }$$
        Such an arrow $\gamma$ is called a \textbf{lift} or a 
        \textbf{solution} to the lifting problem $(u,v)$.
        If such $\gamma$ is unique, i.e. we have:
        $$ \xymatrix{
        a\ar[r]^u\ar[d]^f         & c\ar[d]^g \\
        b\ar[r]^v\ar@{.>}[ru]^{\gamma} & d
        }$$
        Then we say $f$ is \textbf{orthogonal} to $g$, denoted $f \perp g$.
    \end{defi}
    ref \cite{nlab-lift}.

@@

    \paragraph{Fibrations}
    \begin{defi}[Hurewicz fibration]
    \nomenclature{Hurewicz fibration}{\nomrefpage.}
        A map $p$ is called a Hurewicz fibration if it satisfies the homotopy
        lifting property w.r.t to all spaces $X$.
    \end{defi}
    \begin{defi}[Serre fibration]
    \nomenclature{Serre fibration}{\nomrefpage.}
        A map $p$ is called a Serre fibration if it satisfies the homotopy
        lifting property w.r.t to all discs $X$.
    \end{defi}
    Notes: by discs, I think he means closed discs, or closed balls, because he
    also mentioned "equivalently" closed cubes.

    Ref \cite{nlab-homotopy-lifting-property}.
    % Resources: https://ncatlab.org/nlab/show/fibration
    %            https://ncatlab.org/nlab/show/homotopy+lifting+property
    %            https://en.wikipedia.org/wiki/Homotopy_lifting_property
    %            https://en.wikipedia.org/wiki/Fibration
\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{book} D Huybrechts's Introduction to Complex Geometry.
    \bibitem{math.se_1} \href{http://math.stackexchange.com/questions/923957/set-of-almost-complex-structures-on-mathbb-r4-as-two-disjoint-spheres}{set of almost complex structures on $\mathbb{R}^4$ as two disjoint spheres}.
    \bibitem{math.se_2} \href{http://math.stackexchange.com/questions/1356823/does-gln-mathbbc-inject-into-gl2n-mathbbr-for-all-n}{Does $GL(n,\mathbb{C})$ inject into $GL^+(2n, \mathbb{R})$ for all $n$?}
    \bibitem{determinants_of_block_m} John R. Silvester, 
      Determinants of Block Matrices. Available in WebArchive link:
      \url{https://web.archive.org/web/20140505161153/http://www.mth.kcl.ac.uk/~jrs/gazette/blocks.pdf}
    \bibitem{math.se_3} \href{http://math.stackexchange.com/questions/1927487/making-sense-of-frac-mathrmgl2n-mathbbr-mathrmgln-mathbbc/1927518?noredirect=1#comment3957369_1927518}{Making sense of $\frac{\mathrm{GL}(2n,\mathbb{R})}{\mathrm{GL}(n,\mathbb{C})}$}
    \bibitem{math.se_4} \href{http://math.stackexchange.com/questions/57242/similar-matrices-and-field-extensions?noredirect=1&lq=1}
      {Similar matrices and field extensions}
    \bibitem{math.se_set-of-almost} \href{http://math.stackexchange.com/questions/923957/set-of-almost-complex-structures-on-mathbb-r4-as-two-disjoint-spheres}{set of almost complex structures on $\mathbb{R}^4$ as two disjoint spheres}
    \bibitem{nlab-lift}\href{https://ncatlab.org/nlab/show/lift}{nLab Lift}
    \bibitem{nlab-homotopy-lifting-property}\href{https://ncatlab.org/nlab/show/homotopy+lifting+property}{nLab Homotopy lifting property}, \href{https://ncatlab.org/nlab/show/cartesian+product}{nLab Cartesian product}, \href{https://ncatlab.org/nlab/show/interval+object}{nLab Interval object}.
\end{thebibliography}
\printnomenclature

>>>>

    \subsection{Fibration}
    \label{sec:Fibration}
    \paragraph{Lift of morphisms}
    \begin{defi}[Lift of morphisms]
    \nomenclature{Lift of morphisms}{\nomrefpage.}
        The \textbf{lift} of a morphism $f: Y\to B$ along an epimorphism
        \footnote{Epimorphism is the category analogy of surjective
        functions in set theory}. (or more general map) $p:X\to B$ is a
        morphism $\tilde{f}: Y\to X$ such that $f = p\circ\tilde{f}$.
    \end{defi}
    $$
    \xymatrix{
        X\ar[r]^p                    & B \\
        Y\ar[ur]^f\ar[u]^{\tilde{f}} & \\
    }
    $$
    \begin{defi}[Lift property]
    \nomenclature{Lift property}{\nomrefpage.}
        We say that $f$ has a \textbf{left lifting property} w.r.t $g$, or
        equivalently that $g$ has a \textbf{right lifting property} w.r.t
        $f$, if and only if for every commutative diagram below:
        $$
        \xymatrix{
        a\ar[r]^u\ar[d]^f & c\ar[d]^g \\
        b\ar[r]^v         & d
        }
        $$
        there is an arrow $\gamma$, s.t. both triangles in the followinng
        diagram commutes.
        $$ \xymatrix{
        a\ar[r]^u\ar[d]^f         & c\ar[d]^g \\
        b\ar[r]^v\ar[ru]^{\gamma} & d
        }$$
        Such an arrow $\gamma$ is called a \textbf{lift} or a 
        \textbf{solution} to the lifting problem $(u,v)$.
        If such $\gamma$ is unique, i.e. we have:
        $$ \xymatrix{
        a\ar[r]^u\ar[d]^f         & c\ar[d]^g \\
        b\ar[r]^v\ar@{.>}[ru]^{\gamma} & d
        }$$
        Then we say $f$ is \textbf{orthogonal} to $g$, denoted $f \perp g$.
    \end{defi}
    Ref \cite{nlab-lift}.

@@

    \begin{remark}
        (From page 355, Fibration in chapter 11.)

        If one defines the relationship as above, then one will find
        something interesting out of these relations.

        If one defines $h_1:Y\to B$ by $h_1(x)=F(x,0)$, then $f$ is a
        lifting of $h_1$; if one defines $h_2:Y\to B$ by $h_2(x)=F(x,1)$,
        then one sees $F$ actually induces homotopy $h_1\simeq h_2$.
        Similarly, $\sigma$ is also a homotopy $f\simeq \widetilde f$,
        where $\widetilde f=\sigma(x,1)$ is a lifting of 
    \end{remark}
    \paragraph{Fibrations}
    \begin{defi}[Hurewicz fibration]
    \nomenclature{Hurewicz fibration}{\nomrefpage.}
        A map $p$ is called a Hurewicz fibration if it satisfies the homotopy
        lifting property w.r.t to all spaces $X$.
    \end{defi}
    \begin{defi}[Serre fibration]
    \nomenclature{Serre fibration}{\nomrefpage.}
        A map $p$ is called a Serre fibration if it satisfies the homotopy
        lifting property w.r.t to all discs $X$.
    \end{defi}
    Notes: by discs, I think he means closed discs, or closed balls, because he
    also mentioned "equivalently" closed cubes.

    Ref \cite{nlab-homotopy-lifting-property}.
    % Resources: https://ncatlab.org/nlab/show/fibration
    %            https://ncatlab.org/nlab/show/homotopy+lifting+property
    %            https://en.wikipedia.org/wiki/Homotopy_lifting_property
    %            https://en.wikipedia.org/wiki/Fibration

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{book} D Huybrechts's Introduction to Complex Geometry.
    \bibitem{math.se_1} \href{http://math.stackexchange.com/questions/923957/set-of-almost-complex-structures-on-mathbb-r4-as-two-disjoint-spheres}{set of almost complex structures on $\mathbb{R}^4$ as two disjoint spheres}.
    \bibitem{math.se_2} \href{http://math.stackexchange.com/questions/1356823/does-gln-mathbbc-inject-into-gl2n-mathbbr-for-all-n}{Does $GL(n,\mathbb{C})$ inject into $GL^+(2n, \mathbb{R})$ for all $n$?}
    \bibitem{determinants_of_block_m} John R. Silvester, 
      Determinants of Block Matrices. Available in WebArchive link:
      \url{https://web.archive.org/web/20140505161153/http://www.mth.kcl.ac.uk/~jrs/gazette/blocks.pdf}
    \bibitem{math.se_3} \href{http://math.stackexchange.com/questions/1927487/making-sense-of-frac-mathrmgl2n-mathbbr-mathrmgln-mathbbc/1927518?noredirect=1#comment3957369_1927518}{Making sense of $\frac{\mathrm{GL}(2n,\mathbb{R})}{\mathrm{GL}(n,\mathbb{C})}$}
    \bibitem{math.se_4} \href{http://math.stackexchange.com/questions/57242/similar-matrices-and-field-extensions?noredirect=1&lq=1}
      {Similar matrices and field extensions}
    \bibitem{math.se_set-of-almost} \href{http://math.stackexchange.com/questions/923957/set-of-almost-complex-structures-on-mathbb-r4-as-two-disjoint-spheres}{set of almost complex structures on $\mathbb{R}^4$ as two disjoint spheres}
    \bibitem{nlab-lift}\href{https://ncatlab.org/nlab/show/lift}{nLab Lift}
    \bibitem{nlab-homotopy-lifting-property}\href{https://ncatlab.org/nlab/show/homotopy+lifting+property}{nLab Homotopy lifting property}, \href{https://ncatlab.org/nlab/show/cartesian+product}{nLab Cartesian product}, \href{https://ncatlab.org/nlab/show/interval+object}{nLab Interval object}.
    \bibitem{GTM119} Jossph J. Rotman. An Introduction to Algebraic
    Topology. First Edition. GTM 119. Springer.
\end{thebibliography}
\printnomenclature

### 2016-10-10 Lecture note - Group Theory in Physics/group theory in physics.tex
New class notes. A fews new lines in fibration (Complex Geometry)

% Following is for theorems etc environments
% http://tex.stackexchange.com/questions/45817/theorem-definition-lemma-problem-numbering && https://en.wikibooks.org/wiki/LaTeX/Theorems
\usepackage{amsthm}
\newtheorem{defi}{Definition}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{coro}{Corollary}[section]
\theoremstyle{definition}
\newtheorem{ex}{Example}[section]

% A list of nomenclatures.
\usepackage{nomencl}
\makenomenclature

@@

\begin{defi}[$C_N,\bar{C}_N$]
\nomenclature{$C_N,\bar{C}_N$}{\nomrefpage.}
    $\bar{C}_N=C_N*\sigma$
\end{defi}
\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    %\bibitem{book} 
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{math.se_1_lenz_vector} \href{physics.stackexchange.com/questions/18088/what-symmetry-causes-the-runge-lenz-vector-to-be-conserved}{What symmetry causes the Runge-Lenz vector to be conserved?}
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

% Following is for theorems etc environments
% http://tex.stackexchange.com/questions/45817/theorem-definition-lemma-problem-numbering && https://en.wikibooks.org/wiki/LaTeX/Theorems
\usepackage{amsthm}
\newtheorem{defi}{Definition}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{coro}{Corollary}[section]
\theoremstyle{definition}
\newtheorem{ex}{Example}[section]
\newtheorem{fact}{Fact}[section]

\usepackage{CJKutf8} % For chinese characters.
% A list of nomenclatures.
\usepackage{nomencl}
\makenomenclature

@@

\begin{CJK}{UTF8}{gbsn}
固有转动是指的那些 $det(M)>0$ 的转动. 用$C_n$来表示他们.

Also, 周期 of $R$ is just $\braket{R}$.
\end{CJK}

Let $\sigma$\nomenclature{$\sigma$}{\nomrefpage.} for spatial reflection.
\begin{defi}[$C_N,\bar{C}_N$]
\nomenclature{$C_N,\bar{C}_N$}{\nomrefpage.}
    $\bar{C}_N=C_N*\sigma$
\end{defi}
\section{20161010}
\label{sec:20161010}
Introducing to various groups:$S_4, V_4$, $D_3$, all omited.
(\textbf{pp.22-23 of \cite{book}})

$D_n$ group. See pp. 25-26 of the book \cite{book}.
Note that here the $n$ refers to the $n$-polygon, not that the group is
of order $n$. For the mathematicians, they might be comfortable with
$D_n$ means the dihedral group of order $n$, but is actually the group of
symmetries of $n/2$-polygon.

\begin{defi}[Subgroup]
\nomenclature{Subgroup}{\nomrefpage.}
    Omitted.
\end{defi}
\begin{fact}
    One only has to check the closeness for determining a subgroup, if it
    is of finite order.

    However, for group of infinite order, one has to check the existance
    of unit and inverse elements.
\end{fact}
Examples of subgroup (\textbf{pp.26 of \cite{book}})

Noteworth:$C_6$ has three copies of $D_2$, this can be intuitively 
guessed by the fact that a hexago has three rectanle in it.

\begin{defi}[Coset]
\nomenclature{Coset}{\nomrefpage.}
    Omitted.
\end{defi}
Properties of coset (omitted).

\begin{defi}[Index of subgroup]
\nomenclature{Index of subgroup}{\nomrefpage.}
    Omitted.
\end{defi}

\begin{prop}
    Two elements $R$ and $T$ belongs to the same coset $kH$, if and only
    if $R^{-1}T\in H$.
\end{prop}
\begin{proof}
    Omitted.
\end{proof}
\begin{defi}[Normal/Invariant subgroup]
\nomenclature{Normal/Invariant subgroup}{\nomrefpage.}
    A subgroupNormal/Invariant subgrouproup(also invariant), if and only
    if for any $x\in G$, we have $xH = Hx$.
\end{defi}
\begin{fact}
    If $H$ has index 2, then it must be normal/invariant. This is
    obvious.
\end{fact}
\begin{defi}[Quotient]
\nomenclature{Quotient}{\nomrefpage.}
    Omitted.
\end{defi}
Note that quotient group (a.k.a. factor group) is only defined for a normal subgroup.

\begin{ex}
    $D_3$ (Using the multiplication table). Omitted because it is too
    complex to be typed down here.
\end{ex}

\begin{defi}[Conjugate]
\nomenclature{Conjugate}{\nomrefpage.}
    If exists $S\in G$, s.t. $R' = S^{-1}RS$, then we say $R'$ is
    conjugate to $R$.
\end{defi}
see (\textbf{pp.28-30 of \cite{book}})
This is clearly an equivalence relationship. By this we can define
conjugate class, denoted by $C=\{R_1,\cdots\}$, then we have the
characterization $C = \{ s^{-1} R_i s|\forall s\in G\}$, for any
$R_i$. We then have the following facts (all are obvious):
\begin{fact}$ $

    \begin{enumerate}
        \item The unit class formed just by the unit element.
        \item The inverse class formed by just all the inverse element.
            $C^{-1} = \{ R_i^{-1}\}$. If $C=C^{-1}$, then $C$ is called
            self-inverse.
        \item The order of elements in a class is just the same.
        \item For $\forall T,S\in G$, $TS$ and $ST$ are conjugate to each
            other. This means that all elements symmetric on the
            multiplication table is conjugate to each other.
        \item For two elements $R$, $R'$ conjugate to each other, both
            can be expressed by the products of two elements in two
            different way. (Isn't this too obvious to mention.)
        \item Let $G$ be a rotation group. Suppose it has an axis of the
            order of $n$, with its operation denoted as $R$,, we can get
            a new axis by the following steps. (Supose we have another
            rotation $S$),

            \begin{enumerate}
                \item  $S^{-1}$, rotate $m$ back to $n$.
                \item  $R$, rotate about n around $2\pi/n$,
                \item  $S$ rotate $n$ to $m$
            \end{enumerate}

            Result: $S^{-1}RS$ rotate around a new axis $m$ about
            $2\pi/n$. So $R'=S^{-1}RS$ and $R$ is calle the equivalent
            axis.

            Also, if $m=-n$, then they are called polar axis to each
            other.

        \item $C_n$, which is an abel group, every element form a
            conjugate class by itself. Specifically, $e$ and $R^{n/2}$
            are self inverse, if $n$ is even.
    \end{enumerate}
\end{fact}
\begin{prop}
    \label{prop:20161010.conjugate_subgroup}
    For an invariant subgroup, then every conjugate element is also
    inside the same invariant subgroup. This shows that an invariant
    subgroup can be decomposed into a series of sum of conjugate classes.
\end{prop}
\begin{proof}
    If $R\in H$, then we show that $S^{-1}RS\in H$, this is obviously
    since it belongs to $S^{-1}HS$.
\end{proof}

\begin{ex}
    For $D_3: E,D,F,A,B,C$, their orders are respetively
    $1,3,3,2,2,2$. We have the following conjugate classes:
    \begin{enumerate}
        \item $\{ E\}$. Is self inverse.
        \item $\{ D,F\}$. $D$ is conjugate to $F$. We can see
            this physically by looking at rotation from the front
            or the below. This class is also self-inverse.
        \item $\{A,B,C\}$, is clearly a self-inverse conjugate
            class.
    \end{enumerate}
\end{ex}
\begin{ex}[$D_6$]
    Ramiliarize one with the formulae for $D_n$. Hint: use the order of
    elements to find classes of conjugate. Then use the proposition
    \ref{prop:20161010.conjugate_subgroup} to find the subgroups.
    % TODO maybe I can try this when I am free.
\end{ex}

\section{Anchor}
\label{sec:Anchor}
\begin{thebibliography}{1}
    %\bibitem{book} 
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{math.se_1_lenz_vector} \href{physics.stackexchange.com/questions/18088/what-symmetry-causes-the-runge-lenz-vector-to-be-conserved}{What symmetry causes the Runge-Lenz vector to be conserved?}
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
